\documentclass[12pt, letterpaper]{report}

%\usepackage[a4paper, portrait, margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{concmath} % Text Font
\usepackage{amsmath} % Matrix
\usepackage{parskip}
\usepackage{tabularx} % Manage Tables
\usepackage{makecell} % Manage table cells
\usepackage{fancyhdr} % For Headers and Footers
%\usepackage[htt]{hyphenat} % To solve \textit line break problem
\usepackage{hyperref} % For index and web links
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz} % Graph design
\usepackage{pgfplots} % For drawing functions and curves
\usepackage{pdflscape} % Page orientation
\usepackage{float} % Figure positioning
\usepackage{amssymb} % More symbols 
\usepackage{multirow} % single table cell to span multiple rows
\usepackage{xltabular} % for tables spanning multiple pages
\usepackage{siunitx} % for units of measurement
\usepackage{listings}
\lstset{basicstyle=\ttfamily}

\usetikzlibrary{positioning}
\usetikzlibrary{calc}
\usetikzlibrary{arrows.meta}

\graphicspath{ {./img/} }

\title{"Control of Mobile Robot" Project}
\author{Colli Stefano, Pagani Mattia, Panelli Erica}
\date{\today}

\pagestyle{fancy}
\fancyhf{}
\rhead{CMR Project}
\lhead{User Manual}
\rfoot{\thepage}

\begin{document}
	
%\maketitle
\input{user_manual_TITLEPAGE}

\begin{abstract}
The first part of the report contains a description of the original MIT Racecar framework used, on which our modifications and additions are based. Main components, configuration, design files and principles of working are shown. In the central part are inserted schemes and principles of our packages. For each of them there are variables and parameters which have been used, each one with its indications. In addition there is an explanation of the Pacejka plugin to be attached to Gazebo. At the end are shown and described the results obtained in various configurations, even with Pacejka activated. A final appendix is added, with practical references to configuration file inclusions and topics. A brief guide on how to lunch packages is written at the beginning of the report.
\end{abstract}

\tableofcontents

\newpage

%\section{Gazebo Body Parameters}
%
%\begin{itemize}
%	\item $\mu_0$: friction coefficient of direction 1
%	\item $\mu_1$: friction coefficient of direction 2
%	\item fdir: specify direction of $\mu_0$, direction of $\mu_1$ is perpendicular to this parameter
%	\item $k_p$: stiffness between bodies
%	\item $k_d$: dumping between bodies
%\end{itemize}

\chapter{Notes on Installation and Launch}

\section{Installation}

\subsection{Downloading material}

The project is based on the material of original MIT racecar. That's it, we have generated a new ROS environment copying MIT repository packages. In particular the following packages have been downloaded:

\begin{itemize}
	\item ackermann\_msgs
	\item racecar
	\item racecar\_gazebo
\end{itemize} 

They can be found at the link: \url{https://github.com/mit-racecar}

\subsection{Additional packages to be installed}

To be able to compile the project it is necessary to download two internal ROS packages which will be used by the racecar ones. Launch the following commands:

\begin{verbatim}
sudo apt install ros-noetic-ros-control
sudo apt install ros-noetic-ros-controllers
\end{verbatim}

Also Python rospkg should be installed to run bag reading scripts.

\noindent Otherwise an error will be thrown when \verb|catkin_make| command is called.

\subsection{Additional modifications}

In some cases, to avoid conflicts, it's required to change Python environment to version 3 in each file of the original packages.
For example, if Python environment is set to 3, modifications are needed for \verb|joy_teleop.py| file:

\begin{itemize}
	\item Row 277: replace ',' with 'as'
	\item Row 282: replace \verb|iteritems| with \verb|items|
\end{itemize}

These modifications are applied in our version of the libraries, so it's not necessary to update them. We have linked original MIT repository sub-modules updating them.

\section{Launch}

\subsection{Original Project}

In order to launch original project, once it's compiled following ROS guide, following steps should be followed.

There are different configurations which can be used for launching.

\subsubsection{Kinematic linearizer with Gazebo Physics}

\begin{itemize}
	\item roscore
	\item roslaunch src/mit\_racecar/racecar\_gazebo/racecar\_gazebo/launch/racecar.launch [roslaunch racecar\_gazebo racecar.launch]
	\item roslaunch src/car\_kinematic\_linearizer/launch/car\_kin\_linearizer.launch [roslaunch car\_kinematic\_linearizer car\_kin\_linearizer.launch]
	\item roslaunch src/trajectory\_traker/launch/trajectory\_tracker.launch [roslaunch trajectory\_traker trajectory\_traker.launch]
	\item rosbag record -O KinGazebo.bag --duration=1m /reference\_trajectory /vesc/odom /long\_pub/right\_front /lat\_pub/right\_front /fx\_pub/right\_front /fy\_pub/right\_front
\end{itemize}

\subsubsection{Kinematic linearizer with ODE Physics}

\begin{itemize}
	\item roscore
	\item roslaunch ode\_simulator ode\_simulator.launch
	\item roslaunch src/car\_kinematic\_linearizer/launch/car\_kin\_linearizer.launch [roslaunch car\_kinematic\_linearizer car\_kin\_linearizer.launch]
	\item roslaunch src/trajectory\_traker/launch/trajectory\_tracker.launch [roslaunch trajectory\_traker trajectory\_traker.launch]
	\item rosbag record -O KinODE.bag --duration=1m /reference\_trajectory /vesc/odom /long\_pub/right\_front /lat\_pub/right\_front /fx\_pub/right\_front /fy\_pub/right\_front
\end{itemize}

To run using Pacejka it is necessary to set the environment variable PACEJKA=1.

\subsubsection{Kinematic linearizer with Gazebo Physics and Pacejka for wheels}

Same commands of Kinematic linearizer with Gazebo Physics but with Pacejka active.

If there are no errors the user should be able to see the racecar in a Gazebo environment moving depending on chosen trajectory.

\chapter{General Project Structure}

\section{Catkin Workspace Directories}

\subsection{Original MIT Racecar Packages}

\begin{center}
	\begin{tabularx}{\textwidth}{
			| >{\raggedright\arraybackslash}X
			| >{\arraybackslash}X |
		}
		\hline
		ackermann\_cmd\_mux (racecar folder) & ... \\
		\hline
		ackermann\_msgs & Contains definitions of \textbf{AckermannDrive} and \textbf{AckermannDriveStamped} messages, used by the racecar to compute movements. \\
		\hline
		racecar (racecar folder) & Directory which contains \\
		\hline
		racecar\_control (racecar\_gazebo folder) & Contains launch files to load controllers used to manage the motors of the racecar. Also load nodes which dispatch messages to controllers. \\
		\hline
		racecar\_description (racecar\_gazebo folder) & Contains a description of the racecar, in terms of models, meshes ecc... It will be used by Gazebo to represent it. \\
		\hline
		racecar\_gazebo (racecar\_gazebo folder) & Mainly contains launch scripts used to load all necessary nodes, worlds and other components to open a Gazebo instance with a controllable car. \\
		\hline
	\end{tabularx}
\end{center}

\subsection{Added Packages}

\begin{center}
	\begin{tabularx}{\textwidth}{
			| >{\raggedright\arraybackslash}X
			| >{\raggedright\arraybackslash}X |
		}
		\hline
		car\_control & Contains node which performs the linearization of the nonlinear bycicle \textbf{dynamic} model. It' receives desired velocities from trajectory tracker and sends Ackermann commands to the racecar. \\
		\hline
		car\_kinematic\_control & Contains node which performs the exact linearization of the nonlinear bycicle \textbf{kinematic} model. It' receives desired velocities from trajectory tracker and sends Ackermann commands to the racecar. \\
		\hline
		trajectory\_tracker & Generates (or receives in input) a desired trajectory and actual car positions, than compute desired velocities to be sent to controllers. \\
		\hline
	\end{tabularx}
\end{center}

\subsection{Added Plugin}

\begin{center}
	\begin{tabularx}{\textwidth}{
			| >{\raggedright\arraybackslash}X
			| >{\raggedright\arraybackslash}X |
		}
		\hline
		gazebo\_ros\_pacejka & Plugin used to replace inner wheel friction phisical model of ROS with a custom one. \\
		\hline
	\end{tabularx}
\end{center}


\newpage
\chapter{MIT Racecar Model}
\input{mit_racecar.tex}

\newpage
\chapter{(Our) System Description}

\section{Scheme of the whole system}

\begin{tikzpicture}
\node[
	draw,
	fill=BlueGreen,
	minimum width=2cm,
	minimum height=1.5cm
] (tracker) {Trajectory Tracker};

\node[
	draw,
	fill=BlueGreen,
	minimum width=2cm,
	minimum height=1.5cm,
	right=4cm of tracker
] (refVelocity) {Compute Ref. Velocity};

\node[
	draw,
	fill=Green,
	minimum width=2cm,
	minimum height=1.5cm,
	below=3cm of tracker
] (model) {Model MIT/ODE};

\node[
	draw,
	fill=Yellow,
	minimum width=2cm,
	minimum height=1.5cm,
	below=3cm of refVelocity
] (controller) {Linearizer};

\draw[-{Triangle[scale=2]}]  (-3, +1/3) --  ($(tracker.west) + (0, +1/3)$)
	node[midway, above]{$x_{ref}$};
	
\draw[-{Triangle[scale=2]}]  (-3, -1/3) --  ($(tracker.west) + (0, -1/3)$)
	node[midway, above]{$y_{ref}$};

% ---

\draw[-{Triangle[scale=2]}] ($(tracker.east) + (0, +1/3)$) -- ($(refVelocity.west) + (0, +1/3)$)
	node[midway, above]{$x_{current}$};
	
\draw[-{Triangle[scale=2]}] ($(tracker.east) + (0, -1/3)$) -- ($(refVelocity.west) + (0, -1/3)$)
	node[midway, above]{$y_{current}$};
	
% ---
	
\draw[-{Triangle[scale=2]}] ($(refVelocity.south) + (-1/2, 0)$) -- ($(controller.north) + (-1/2, 0)$)
	node[midway, left]{$V_{Xp}$};
	
\draw[-{Triangle[scale=2]}] ($(refVelocity.south) + (+1/2, 0)$) -- ($(controller.north) + (+1/2, 0)$)
	node[midway, left]{$V_{Yp}$};

% ---

\draw[-{Triangle[scale=2]}] ($(controller.west) + (0, +1/2)$) -- ($(model.east) + (0, +1/2)$)
	node[midway, above]{$\phi$};
	
\draw[-{Triangle[scale=2]}] ($(controller.west) + (0, 0)$) -- ($(model.east) + (0, 0)$)
	node[midway, above]{$V$};
	
\draw[-{Triangle[scale=2]}] ($(controller.west) + (0, -1/2)$) -- ($(model.east) + (0, -1/2)$)
	node[midway, above]{$\omega$};

% ---

\draw[-{Triangle[scale=2]}] ($(model.north) + (-1/2, 0)$) -- ($(tracker.south) + (-1/2, 0)$)
	node[near end, left]{$x_{in}$}
	node[near start, left]{$x_{out}$};
	
\draw[-{Triangle[scale=2]}] ($(model.north) + (+1/2, 0)$) -- ($(tracker.south) + (+1/2, 0)$)
	node[near end, left]{$y_{in}$}
	node[near start, left]{$y_{out}$};
	
\draw[-{Triangle[scale=2]}] (model.south) |- ++ (0, -1) -| (controller.south)
	node[near start, above]{$\theta_{out}$};

\end{tikzpicture}

\vspace{1cm}

\begin{center}
	\begin{tabularx}{\textwidth}{|l|X|}
		\hline
		\textbf{Symbol} & \textbf{Meaning} \\
		\hline
		$x_{ref}, y_{ref}$ & Ref. position of the trajectory \\
		\hline
		$V_{Xp}, V_{Yp}$ & Required velocities of the point. They will be imposed by controller\\
		\hline
		$\phi$ & Steer degree of rotation \\
		\hline
		$V$ & Vector velocity \\
		\hline
		$\omega$ & Steer speed of rotation \\
		\hline
		$\theta_{out}$ & Car pose: rotation around center axis\\
		\hline
		$x_{out}, y_{out}$ & Car pose: x, y \\
		\hline
	\end{tabularx}
\end{center}

\vspace{1cm}

Note that "trajectory tracker" generated trajectories are hard-coded, even if $x_{ref}$ and $y_{ref}$ are shown as input parameters. The user can select the trajectory using YAML configuration file (which will be explained in the relative section).

\section{Topics}

\subsection{Scheme of topic publications/subscriptions}

\subsubsection{ROS Friction Model}

\begin{tikzpicture}
	
\node[
	draw,
	fill=BlueGreen,
	minimum width=2cm,
	minimum height=1.5cm
] (tracker) {Trajectory Tracker};

\node[
	draw,
	fill=White,
	minimum width=2cm,
	minimum height=1.5cm,
	right=4cm of tracker
] (virtualVel) {/virtual\_velocities};

\node[
	draw,
	fill=White,
	minimum width=2cm,
	minimum height=1.5cm,
	below=3cm of tracker
] (odom) {/vesc/odom};

\node[
	draw,
	fill=Yellow,
	minimum width=2cm,
	minimum height=1.5cm,
	below=3cm of virtualVel
] (controller) {Linearizer};

\node[
	draw,
	fill=Green,
	minimum width=2cm,
	minimum height=1.5cm,
	below=3cm of odom
] (model) {Model MIT/ODE};

\node[
	draw,
	fill=White,
	minimum width=2cm,
	minimum height=1.5cm,
	below=3cm of controller
] (ackermanCmd) {/vesc/ackermann\_cmd\_mux/input/navgation};

% ---

\draw[-{Triangle[scale=2]}] (tracker.east) -- (virtualVel.west)
	node[midway, above]{PUB};
	
\draw[-{Triangle[scale=2]}] (virtualVel.south) -- (controller.north)
	node[midway, right]{};
	
\draw[-{Triangle[scale=2]}] (controller.south) -- (ackermanCmd.north)
	node[midway, right]{PUB};
	
\draw[-{Triangle[scale=2]}] (ackermanCmd.west) -- (model.east)
	node[midway, above]{};
	
\draw[-{Triangle[scale=2]}] (model.north) -- (odom.south)
	node[midway, right]{PUB};
	
\draw[-{Triangle[scale=2]}] (odom.north) -- (tracker.south)
	node[midway, right]{};
	
\draw[-{Triangle[scale=2]}] (odom.east) -- (controller.west)
	node[midway, above]{};

\end{tikzpicture}


\subsection{Topics meaning}

\subsubsection{Common Topics}

\begin{center}
	\begin{tabularx}{\textwidth}{
			| >{\raggedright\arraybackslash}X
			| >{\raggedright\arraybackslash}X |
		}
		\hline
		/virtual\_velocities & Used by "trajectory tracker" to publish desired velocity components. These are read by controller in order to perform linearization and compute instructions for the model. \\
		\hline
		\makecell[lt]{/vesc/ackermann\_cmd\_mux \\ /input/navgation} & Contains AckermannDriveStamped messages sent by controller. These messages contains information for the racecar, about velocity and steering. \\
		\hline
		/vesc/odom & The model uses this topic to publish odometry information of the racecar (position and orientation). These data are used both by tracker and controller. The first one compute differences between actual car position and desired position imposed by trajectory. The last one reads z-axis orientation useful to perform linearization. \\
		\hline
	\end{tabularx}
\end{center}

There is another topic in which "trajectory tracker" publish, the \textit{/reference\_trajectory}. This is used to read trajectory information to perform debug and register data for analysis.

\vspace{1cm}

\begin{tikzpicture}
	
\node[
	draw,
	fill=BlueGreen,
	minimum width=2cm,
	minimum height=1.5cm
] (tracker) {Trajectory Tracker};

\node[
	draw,
	fill=White,
	minimum width=2cm,
	minimum height=1.5cm,
	right=4cm of tracker
] (reference) {/reference\_trajectory};

\draw[-{Triangle[scale=2]}] (tracker.east) -- (reference.west)
	node[midway, above]{PUB};

\end{tikzpicture}

%\subsubsection{Specific Topics}
%
%\begin{center}
%	\begin{tabularx}{\textwidth}{
%			| >{\raggedright\arraybackslash}X
%			| >{\raggedright\arraybackslash}X |
%		}
%		\hline
%		\makecell[lt]{/vesc/low\_level/ \\ ackermann\_cmd\_mux/output} & xyz \\
%		\hline
%	\end{tabularx}
%\end{center}

\chapter{Detailed Package Description}

\section{Package car\_dynamic\_linearizer}
\input{car_dynamic_linearizer}

\newpage

\section{Package car\_kinematic\_linearizer}
\input{car_kinematic_linearizer}

\newpage

\section{Package trajectoy\_tracker}
\input{trajectory_tracker}

\newpage

\section{Package ode\_simulator}
\input{ode_simulator}

\newpage

\section{Plugin gazebo\_ros\_pacejka}
\input{pacejka_plugin}

\newpage

\chapter{Python Scripts}
\input{python_scripts.tex}

\newpage

\chapter{Conclusions}
\input{conclusions}

\appendix

\chapter{launch package inclusion}
	
	\usetikzlibrary{arrows.meta}
\begin{figure}[H]
	\makebox[\textwidth][c] 
	{
	\begin{tikzpicture}[
		scale=0.7,
		level/.style={level distance=10em, sibling distance=21em},
		every node/.style = {
			shape=rectangle, 
			solid,
			draw, 
			align=center,
			top color=white, 
			bottom color=green!20
		},
		norm/.style={edge from parent/.style={-{Latex[scale=2]}, solid, draw}},
		emph/.style={edge from parent/.style={-{Latex[scale=2]}, dashed, draw}}
		]
		\node {\textbf{PKG: racecar\_control} \\ gazebo\_sym\_joy.launch \\ <perform remap>}
		child [norm] { node {\textbf{PKG: racecar\_gazebo} \\ racecar\_tunnel.launch} 
			child [norm] { node {\textbf{PKG: racecar\_gazebo} \\ racecar.launch} 
				child [norm] { node {\textbf{PKG: racecar\_control} \\ racecar\_control.launch \\ <perform remap>} 
					child [norm] { node {\textbf{PKG: racecar\_control} \\ servo\_commands.py \\ <perform subs>} edge from parent node {\textit{node}} }
					child [norm] { node {\textbf{PKG: racecar\_gazebo} \\ gazebo\_odometry.py} edge from parent node {\textit{node}} }
					edge from parent node {\textit{include}}
				}
				child [norm] { node {\textbf{PKG: racecar} \\ mux.launch} edge from parent node {\textit{include}} }
				edge from parent node {\textit{include}}
			} 
			edge from parent node {\textit{include}}
		}
		child [norm] { node {\textbf{PKG: racecar\_control} \\ teleop.launch} 
			child [emph] { node {\textbf{PKG: racecar} \\ joy\_teleop.launch.xml} 
				child [norm] { node {\textbf{PKG: racecar} \\ joy\_teleop.py} edge from parent node {\textit{node}} }
				edge from parent node {\textit{include}}
			}
			edge from parent node {\textit{include}}
		};
	\end{tikzpicture}
	}
\end{figure}
	
%	\begin{tikzpicture}[
%		sibling distance=18em,
%		every node/.style = {
%			shape=rectangle, 
%			draw, 
%			align=center,
%			top color=white, 
%			bottom color=green!20
%		},
%		edge from parent/.style= {
%			draw,
%			-latex
%		}
%		]
%		\node {\textbf{PKG: racecar\_control} \\ keyboard\_teleop.py \\ <perform publish>};
%	\end{tikzpicture}

\begin{center}
	\begin{tabularx}{\linewidth}{|X|X|X|}
		\hline
		\textbf{Package} & \textbf{File} & \textbf{Remap} \\
		\hline
		racecar & joy\_teleop.launch.xml & (none) \\
		\hline
		racecar & joy\_teleop.py & (none) \\
		\hline
		racecar & mux.launch & (none) \\
		\hline
		racecar\_control & gazebo\_sim\_joy.launch & \parbox[t]{5cm}{\raggedright \textcolor{Green}{REMAP \\ /ackermann\_cmd\_ \\ mux/input/teleop \\ TO \\ /racecar/ackermann\_ \\ cmd\_mux/input/teleop}} \\
		\hline
		racecar\_control & teleop.launch & (none) \\
		\hline
		racecar\_control & racecar\_control.launch & \parbox[t]{5cm}{\raggedright \textcolor{Green}{REMAP \\ /racecar/ack/output \\ TO \\ /vesc/low\_level/ \\ ack/output}} \\
		\hline
		racecar\_control & servo\_commands.py & \parbox[t]{5cm}{\raggedright \textcolor{Green}{SUBSCRIBE \\ /racecar/ackermann\_ \\ cmd\_mux/output}} \\
		\hline
		racecar\_control & keybpard\_teleop.py & \parbox[t]{5cm}{\raggedright \textcolor{Green}{PUBLISH \\ /vesc/achermann\_ \\ cmd\_mux/input/teleop}} \\
		\hline
		racecar\_gazebo & racecar\_tunnel.launch & (none) \\
		\hline
		racecar\_gazebo & racecar.launch & (none) \\
		\hline
		racecar\_gazebo & gazebo\_odometry.py & (none) \\
		\hline
	\end{tabularx}
\end{center}

\end{document}